---
title: "Gov50 Final Project"
subtitle: "Comparing and Analyzing the Rates of Success in Competitive High School Debate Between Private and Public Schools"
author: "Omar Sotelo and Grace Kim"
output:
  pdf_document: default
  html_document: default
---

```{r echo=FALSE}
library(magrittr)
library(stringr)
library(rvest)


export <- function(tournament, link){
x <- read_html(link) %>% html_nodes("table") %>% html_table() %>% as.data.frame()
}

exportberkone <- function(tournament, link){
x <- read_html(link) %>% html_nodes("table:nth-of-type(1)") %>% html_table() %>% as.data.frame()
}
exportberktwo <- function(tournament, link){
x <- read_html(link) %>% html_nodes("table:nth-of-type(2)") %>% html_table() %>% as.data.frame()
}
exportberkthree <- function(tournament, link){
x <- read_html(link) %>% html_nodes("table:nth-of-type(3)") %>% html_table() %>% as.data.frame()
}

## I want all my sheets to have the same format, all I need from the following datasets are, name, school, state, preliminary speaks, and Z score.

applevalley <- export(applevalley, 
                      "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16856&result_id=141519")

berkeleyone <- exportberkone(berkeley,
                   "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16917&result_id=166166")

berkeleytwo <- exportberktwo(berkeley,
                   "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16917&result_id=166166")

berkeleythree <- exportberkthree(berkeley,
                   "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16917&result_id=166166")

bronx <- export(bronx,
                "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=15710&result_id=136556")
  
emory <- export(emory,
                "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16030&result_id=160506")

glenbrooks <- export(glenbrooks,
                     "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=14991&result_id=146747")

greenhill <- export(greenhill,
                    "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16788&result_id=132929")

harvard <- export(harvard,
                  "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16776&result_id=166103")

harvard.westlake <- export(harvard.westlake,
                           "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=18127&result_id=156949")

heartoftexas <- export(heartoftexas,
                       "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=16679&result_id=170440")

midamerica <- export(midamerica,
                     "https://www.tabroom.com/index/tourn/results/event_results.mhtml?tourn_id=20873&result_id=188728")



# Cleaning up Berkeley
berkeleyone <- berkeleyone[c('First', 'Last', 'Entry', 'State', 'Z2', 'Pts..1HL')]
berkeleytwo <- berkeleytwo[c('First', 'Last', 'Entry', 'State', 'Z2', 'Pts..1HL')]
berkeleythree <- berkeleythree[c('First', 'Last', 'Entry', 'State', 'Z2', 'Pts..1HL')]
berkeley <- rbind(berkeleyone, berkeleytwo, berkeleythree)


## Adjust column names so the Z scores are all just called Z
## Emory, Glenbrooks, and Harvard already call it Z
# Ones called Z2
# AppleValley
applevalley$Z <- applevalley$Z2

# Berkeley
berkeley$Z <- berkeley$Z2

# Greenhill
greenhill$Z <- greenhill$Z2

# Harvard-Westlake
harvard.westlake$Z <- harvard.westlake$Z2

# Mid America
midamerica$Z <- midamerica$Z2

## Z1Pm
# Bronx
bronx$Z <- bronx$Z1Pm

## Z..1HL
# Heart of Texas
heartoftexas$Z <- heartoftexas$Z..1HL


## I also need to adjust the names for the regular speaks
# Bronx
bronx$Pts..1HL <- bronx$PtsPm..1HL

# Harvard Westlake doesn't release their normal speaks in the normal chart but we can calculate them using the ballots column
harvard.westlake[c('r1', 'r2', 'r3', 'r4', 'r5', 'r6')] <- as.numeric(str_split_fixed(harvard.westlake$Ballots, ',', 6))
for (i in 1:nrow(harvard.westlake)){
  speaks <- c(harvard.westlake$r1[i], harvard.westlake$r2[i], harvard.westlake$r3[i], harvard.westlake$r4[i], harvard.westlake$r5[i], harvard.westlake$r6[i])
  if(is.na(harvard.westlake$r6[i]) == TRUE){
    harvard.westlake$r6[i] <- mean(speaks, na.rm = TRUE)
    speaks <- c(harvard.westlake$r1[i], harvard.westlake$r2[i], harvard.westlake$r3[i], harvard.westlake$r4[i], harvard.westlake$r5[i], harvard.westlake$r6[i])
  }
  harvard.westlake$Pts..1HL[i] <- sum(speaks, na.rm = TRUE) - min(speaks, na.rm = TRUE) - max(speaks, na.rm = TRUE)
  
}


## I want first and last names to be separated but Bronx combines the names so I will split this column up
bronx[c('First', 'Last')] <- str_split_fixed(bronx$Name, ' ', 2)


## Adjusting Entry Name
# Bronx calls their entries "Code" instead of "Entry"
bronx$Entry <- bronx$Code
# Mid America Cup doesn't have an Entry Column but we can Download another dataset that does from the same tournament and merge the two since this one won't have Z scores
midamerica.entry <- export(midamerica.entry,
                     "https://www.tabroom.com/index/tourn/results/ranked_list.mhtml?event_id=180185&tourn_id=20873")
midamerica$Name <- str_c(midamerica$First, ' ', midamerica$Last)
midamerica <- merge(midamerica, midamerica.entry, by = "Name")
midamerica$School <- midamerica$School.x
midamerica$Entry <- midamerica$School



## Rearrange and Clean Up Data
applevalley <- applevalley[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
berkeley <- berkeley[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
bronx <- bronx[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
emory <- emory[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
glenbrooks <- glenbrooks[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
greenhill <- greenhill [c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
harvard <- harvard[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
harvard.westlake <- harvard.westlake[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
heartoftexas <- heartoftexas[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]
midamerica <- midamerica[c('First', 'Last', 'Entry', 'State', 'Z', 'Pts..1HL')]



## Adjusting Entry Name so it only has school
applevalley$Entry <- substr(applevalley$Entry,1,nchar(applevalley$Entry)-3)
berkeley$Entry <- substr(berkeley$Entry,1,nchar(berkeley$Entry)-3)
bronx$Entry <- substr(bronx$Entry,1,nchar(bronx$Entry)-3)
emory$Entry <- substr(emory$Entry,1,nchar(emory$Entry)-3)
glenbrooks$Entry <- substr(glenbrooks$Entry,1,nchar(glenbrooks$Entry)-3)
greenhill$Entry <- substr(greenhill$Entry,1,nchar(greenhill$Entry)-3)
harvard$Entry <- substr(harvard$Entry,1,nchar(harvard$Entry)-3)
harvard.westlake$Entry <- substr(harvard.westlake$Entry,1,nchar(harvard.westlake$Entry)-3)
heartoftexas$Entry <- substr(heartoftexas$Entry,1,nchar(heartoftexas$Entry)-3)


## Add tournamnet name so we can differentiate before we combine
applevalley$tournament <- "AppleValley"
berkeley$tournament <- "Berkeley"
bronx$tournament <- "Bronx"
emory$tournament <- "Emory"
glenbrooks$tournament <- "Glenbrooks"
greenhill$tournament <- "Greenhill"
harvard$tournament <- "Harvard"
harvard.westlake$tournament <- "HarvardWestlake"
heartoftexas$tournament <- "HeartofTexas"
midamerica$tournament <- "MidAmerica"


## Sorting Schools by Public VS Private
# Make a big list of all the entries
all <- rbind(applevalley, berkeley, bronx, emory, glenbrooks, greenhill, harvard, harvard.westlake, heartoftexas, midamerica)

# Some schools might cause problems because they have a space after their name, we can delete that and recombine data
all$Entry <- trimws(all$Entry)
all$Combined <- str_c(all$Entry," ", all$State)

# To get list of all unique schools 
schools <- unique(all$Combined) 

# To export list of all unique schools with states 
schools <- as.data.frame(schools)

# Only School Names without State
schools <- unique(all$Combined)
schools <- as.data.frame(schools)


## After I got the dataset of unique schools, I manually sorted the schools into private or public because
# there was no efficient way to sort the schools, especially because schools sometimes use a slightly different
# competitive name from their official school name and some students compete independently from their school which
# slightly alters the entry name

# Import Sorted Data Set with School Type
school.type <- read.csv("schooltypes.csv")
school.type$Entry <- school.type$School.Name

# Add School Type to All Data
schooltype <- rep(NA, nrow(all))
for(i in 1:nrow(all)) {
  for(j in 1:nrow(school.type)){
    if(all$Combined[i] == school.type$Combined[j]){
      if(school.type$Private.Public[j] == "Private"){
        schooltype[i] <- "Private"
        break
      }
      else{
        schooltype[i] <- "Public"
      }
    }
  }
}

all$School.Type <- schooltype
all <- all[c('tournament', 'First', 'Last', 'Entry', 'State', 'Z', 'School.Type', 'Pts..1HL')]


write.csv(all, "all.csv")
write.csv(applevalley, "applevalley.csv")
write.csv(berkeley, "berkeley.csv")
write.csv(bronx, "bronx.csv")
write.csv(emory, "emory.csv")
write.csv(glenbrooks, "glenbrooks.csv")
write.csv(greenhill, "greenhill.csv")
write.csv(harvard, "harvard.csv")
write.csv(harvard.westlake, "harvard.westlake.csv")
write.csv(heartoftexas, "heartoftexas.csv")
write.csv(midamerica, "midamerica.csv")


```

## **Introduction:**  
The question we wanted to answer with our project was, what is the difference in competitive success between private and public schools in high school debate? There is a huge disparity in the number of resources offered to students at public and private schools. We think that comparing the competitive outcomes between these two groups could give policy makers key insight about resource disparities and educational inequality. Our hypothesis is that Private schools have statistically significant higher rates of competitive success in high school debate than public schools. This means that we should see debaters from private schools receiving higher speaker scores and z-scores than debaters from public schools.  


**Dataset:** We decided to limit our dataset to LD octas bid tournaments from 2020-2021 for a few reasons. First, we decided to limit the time-frame to 2020-2021 because it would eliminate the possibility of varying results do to a transition to online debate. We also wanted to limit our dataset to just LD debate because it is a 1v1 format versus other forms of debate being 2v2 or 3v3. This allowed us to look at individual competitive success. Finally, we used only octas bid tournaments because these are national tournaments that are the most organized, have the most attendees, most qualified judges, and most impactful for debaters’ competitive success. In order to compile our data, first, we created a scrape function to scrape the data and export it to a dataframe from the website tabroom.com which is a website that hosts debate tournaments online as well as stores the results of past tournaments. Then we cleaned up our data so each dataset would have the same columns and column names, this was very code intensive (over 200 lines of code!). We also manually created a key for each unique school in our dataset because some schools had weird names (ex: independent entries), and there were only 300 unique school names. When we are analyzing our data, we will be utilizing the 1HL Speaker Scores and the Z-scores. For context, in Lincoln-Douglas Debate, students are awarded a speaker score ranging from 0-30 (speaker scores most typically range from 25-30, 25 being very bad and 30 being perfect; anything below 25 is reserved for speeches that had some other problem, like if the speaker said something offensive). The 1HL Speaker Total Score is the sum of the speaker points subtracting the highest and lowest speaks. The Z-score is a score that just sums the speaks but tries to account for judge biases (ie. some judges just naturally give higher or lower speaks) and standardized everyone's score.

**Methodology:** First, we will take initial counts of public vs. private school attendance at tournaments to gauge representation. Next, we will start our initial mean comparisons and to get an initial understanding of our data. Then, we will switch our analysis to be tournament by tournament and round by round. We will then find the difference in round average speaks and find the 95% and 99% confidence intervals for the difference between private and public schools. We will then run these tests again using Z-score to try and account for judge biases.

```{r echo=FALSE}
countstudents <- table(all$School.Type)
table <- prop.table(countstudents)

```

## **Analysis:**
**Initial Count of Students:** We started with an initial count of the number of public vs. private school students to gauge tournament attendance by school type. We found that there are `r countstudents[1]` in attendance from private schools and `r countstudents[2]` in attendance from public schools. There are nearly double the amount of public school students as private school students at these tournaments. While this may seem like there are a lot of public school students at these tournaments, there are proportionally pretty low attendance (much more public students in the US than private and yet private school students make up about 1/3 of tournament attendance)

```{r echo=FALSE}
publicmean_all <- mean(all$Pts..1HL[all$School.Type == "Public"], na.rm = TRUE)
privatemean_all <- mean(all$Pts..1HL[all$School.Type == "Private"], na.rm = TRUE)

```

**Mean of Student Speaks:** This initial mean of student speaks at schools will give us a baseline understanding of our data. When we average the scores for public schools across all tournaments, we get `r publicmean_all`. However, the private school average comes out to be `r privatemean_all`. This initial mean comparison shows us that public schools students perform better on average. This actually goes against our initial hypothesis that private schools will compete better. However, there may be a confounding variables at play such as the number of rounds per tournament. We can try to account for this by calculating the mean speaker score by tournament instead of overall. We will then create a graph to show tournament by tournament averages.

```{r echo=FALSE, fig1, fig.height = 3, fig.width = 5, fig.align = 'center'}

# Means per tournament
publicmean.applevalley <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "AppleValley"])
privatemean.applevalley <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "AppleValley"])

publicmean.berkeley <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "Berkeley"], na.rm = TRUE)
privatemean.berkeley <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "Berkeley"], na.rm = TRUE)

publicmean.bronx <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "Bronx"])
privatemean.bronx <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "Bronx"])

publicmean.emory <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "Emory"])
privatemean.emory <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "Emory"])

publicmean.glenbrooks <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "Glenbrooks"], na.rm = TRUE)
privatemean.glenbrooks <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "Glenbrooks"], na.rm = TRUE)

publicmean.greenhill <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "Greenhill"])
privatemean.greenhill <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "Greenhill"])

publicmean.harvard <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "Harvard"])
privatemean.harvard <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "Harvard"])

publicmean.harvardwestlake <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "HarvardWestlake"])
privatemean.harvardwestlake <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "HarvardWestlake"])

publicmean.heartoftexas <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "HeartofTexas"])
privatemean.heartoftexas <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "HeartofTexas"])

publicmean.midamerica <- mean(all$Pts..1HL[all$School.Type == "Public" & all$tournament == "MidAmerica"])
privatemean.midamerica <- mean(all$Pts..1HL[all$School.Type == "Private" & all$tournament == "MidAmerica"])


# Create Vectors
publicmeans <- c(publicmean.applevalley, publicmean.berkeley, publicmean.bronx, publicmean.emory, publicmean.glenbrooks, publicmean.greenhill, publicmean.harvard, publicmean.harvardwestlake, publicmean.heartoftexas, publicmean.midamerica)
privatemeans <- c(privatemean.applevalley, privatemean.berkeley, privatemean.bronx, privatemean.emory, privatemean.glenbrooks, privatemean.greenhill, privatemean.harvard, privatemean.harvardwestlake, privatemean.heartoftexas, privatemean.midamerica)


# Bind Vectors
means <- rbind(publicmeans, privatemeans)

# Plot

par(mar=c(3,4,3,1))

barplot(means,
        names.arg = c("A", "Bk", "Bx", "E", "Gx", "Gh", "Hv", "Hw", "Ht", "M"),
        col = c("lightskyblue2", "mistyrose2"),
        main = "Mean Speaks by Tournament Public vs. Private",
        ylab = "Mean Speaker Score",
        cex.axis = 1,
        cex.main = 1,
        beside = TRUE)
legend("bottomright", legend = c("Public", "Private"), col = c("lightskyblue2", "mistyrose2"), pch = 19, bg = "white")

```

This graph shows that on average, private schools are competing better at every tournament except Heart of Texas. It also shows us that certain tournaments like Bronx, Glenbrooks, and Harvard all have significantly higher speaker averages than the other tournaments. This is because those three tournaments actually have more one more preliminary round than the other tournaments that gets added into the students' total speaks. In order to be able to compare tournaments to each other without the number of rounds skewing the data, we can find the average speaks per round by tournament instead of looking at the total speaks by tournament.

```{r echo=FALSE, fig2, fig.height = 3, fig.width = 5, fig.align = 'center'}

tournamentrounds <- c(4, 4, 5, 4, 5, 4, 5, 4, 4, 4)
roundaveragepublic <- publicmeans/tournamentrounds
roundaverageprivate <- privatemeans/tournamentrounds
rbr1hl <- rbind(roundaveragepublic, roundaverageprivate)
par(mar=c(3,4,3,1))

barplot(rbr1hl,
        names.arg = c("A", "Bk", "Bx", "E", "Gx", "Gh", "Hv", "Hw", "Ht", "M"),
        col = c("lightskyblue2", "mistyrose2"),
        main = "Mean Round Speaks by Tournament",
        ylab = "Mean Speaker Score",
        ylim = c(28, 29),
        xpd = FALSE,
        cex.axis = 1,
        cex.main = 1,
        beside = TRUE)
legend("bottomright", legend = c("Public", "Private"), col = c("lightskyblue2", "mistyrose2"), pch = 19, bg = "white")

```


```{r echo=FALSE}
all$average <- ifelse(all$tournament == "Bronx" | all$tournament == "Glenbrooks" | all$tournament == "Harvard", all$Pts..1HL/5, all$Pts..1HL/4)

private <- subset(all, School.Type == "Private")
public <- subset(all, School.Type == "Public")
```

Now that we know that the tournament rounds is a confounding variable, we can account for it by using the round average rather than total speaks in further calculations. We should also find the round average for each individual rather than the overall tournament round averages so that we do not weigh all tournaments equal to each other (it is important that we do not weigh these tournaments equal to each other because certain tournaments have higher attendance than others and the proportions for attendance vary as well - we do not want tournament size to affect our results, rather we want to look at individual performance between public and private school students). We can do this by subsetting students by public vs. private rather than by tournament. We will also do a means comparison of round averages between all public school students and all private school students. The initial mean for public schools was `r mean(public$average)`, while the initial mean for private schools was `r mean(private$average)`.  

```{r echo=FALSE}
# X Bar
xbar_diff <- mean(private$average) - mean(public$average)

# Standard Error
se_diff <- sqrt((var(private$average)/length(private$average)) + (var(public$average)/length(public$average)))
```

```{r echo=FALSE}
# 95% Confidence Interval
m95 <- c(xbar_diff - qnorm(0.975) * se_diff, xbar_diff + qnorm(0.975) * se_diff)

# 99% Confidence Interval
m99 <- c(xbar_diff - qnorm(0.995) * se_diff, xbar_diff + qnorm(0.995) * se_diff)
```
**Confidence Interval Testing:** Next, we want to know if this difference in public vs. private school student averages was statistically significant. We can find this out by calculating a 95% and a 99% difference in means confidence interval and testing to see if 0 is a value within either confidence interval.  

95% Confidence Interval: `r m95` and 99% Confidence Interval: `r m99`  

Since 0 is not a value in either the 95% or the 99% confidence interval we can say that the difference in performance between public and private schools is significant.

However, it is possible that judge preferences or judge biases can effect the speaker scores (ie. some judges just give lower speaks while some judges just give higher speaks). Thus, we can use the Z scores to try and account for judge bias. We will repeat all the same analysis that we did for 1HL Speaker Scores but use Z Scores instead.

```{r echo=FALSE}
z.public.mean <- mean(all$Z[all$School.Type == "Public"], na.rm = TRUE)
z.private.mean <- mean(all$Z[all$School.Type == "Private"], na.rm = TRUE)
```
**Z-Scores**
Public schools Z mean: `r z.public.mean`, Private schools Z mean: `r z.private.mean`

Again, the initial mean goes against our hypothesis because it shows that the public school mean Z score is higher than the private school mean. We will run the same tournament by tournament analysis.

```{r echo=FALSE, fig3, fig.height = 3.2, fig.width = 5, fig.align = 'center'}
# Means per tournament
publiczmean.applevalley <- mean(all$Z[all$School.Type == "Public" & all$tournament == "AppleValley"])
privatezmean.applevalley <- mean(all$Z[all$School.Type == "Private" & all$tournament == "AppleValley"])


publiczmean.berkeley <- mean(all$Z[all$School.Type == "Public" & all$tournament == "Berkeley"], na.rm = TRUE)
privatezmean.berkeley <- mean(all$Z[all$School.Type == "Private" & all$tournament == "Berkeley"], na.rm = TRUE)

publiczmean.bronx <- mean(all$Z[all$School.Type == "Public" & all$tournament == "Bronx"])
privatezmean.bronx <- mean(all$Z[all$School.Type == "Private" & all$tournament == "Bronx"])

publiczmean.emory <- mean(all$Z[all$School.Type == "Public" & all$tournament == "Emory"])
privatezmean.emory <- mean(all$Z[all$School.Type == "Private" & all$tournament == "Emory"])

publiczmean.glenbrooks <- mean(all$Z[all$School.Type == "Public" & all$tournament == "Glenbrooks"], na.rm = TRUE)
privatezmean.glenbrooks <- mean(all$Z[all$School.Type == "Private" & all$tournament == "Glenbrooks"], na.rm = TRUE)

publiczmean.greenhill <- mean(all$Z[all$School.Type == "Public" & all$tournament == "Greenhill"])
privatezmean.greenhill <- mean(all$Z[all$School.Type == "Private" & all$tournament == "Greenhill"])

publiczmean.harvard <- mean(all$Z[all$School.Type == "Public" & all$tournament == "Harvard"])
privatezmean.harvard <- mean(all$Z[all$School.Type == "Private" & all$tournament == "Harvard"])

publiczmean.harvardwestlake <- mean(all$Z[all$School.Type == "Public" & all$tournament == "HarvardWestlake"])
privatezmean.harvardwestlake <- mean(all$Z[all$School.Type == "Private" & all$tournament == "HarvardWestlake"])

publiczmean.heartoftexas <- mean(all$Z[all$School.Type == "Public" & all$tournament == "HeartofTexas"])
privatezmean.heartoftexas <- mean(all$Z[all$School.Type == "Private" & all$tournament == "HeartofTexas"])

publiczmean.midamerica <- mean(all$Z[all$School.Type == "Public" & all$tournament == "MidAmerica"])
privatezmean.midamerica <- mean(all$Z[all$School.Type == "Private" & all$tournament == "MidAmerica"])


# Create Vectors
publiczmeans <- c(publiczmean.applevalley, publiczmean.berkeley, publiczmean.bronx, publiczmean.emory, publiczmean.glenbrooks, publiczmean.greenhill, publiczmean.harvard, publiczmean.harvardwestlake, publiczmean.heartoftexas, publiczmean.midamerica)
privatezmeans <- c(privatezmean.applevalley, privatezmean.berkeley, privatezmean.bronx, privatezmean.emory, privatezmean.glenbrooks, privatezmean.greenhill, privatezmean.harvard, privatezmean.harvardwestlake, privatezmean.heartoftexas, privatezmean.midamerica)


# Bind Vectors
zmeans <- rbind(publiczmeans, privatezmeans)

# Plot

par(mar=c(3,4,3,1))

barplot(zmeans,
        names.arg = c("A", "Bk", "Bx", "E", "Gx", "Gh", "Hv", "Hw", "Ht", "M"),
        col = c("lightskyblue2", "mistyrose2"),
        main = "Mean Z Score by Tournament for Public vs. Private",
        ylab = "Mean Z Score",
        cex.axis = 1,
        cex.main = 1,
        beside = TRUE)
legend("bottomright", legend = c("Public", "Private"), col = c("lightskyblue2", "mistyrose2"), pch = 19, bg = "white")

```


```{r echo=FALSE, fig4, fig.height = 3.7, fig.width = 5.5, fig.align = 'center'}
tournamentrounds <- c(6, 6, 7, 6, 7, 6, 7, 6, 6, 6)
zroundaveragepublic <- publiczmeans/tournamentrounds
zroundaverageprivate <- privatezmeans/tournamentrounds
zrbr <- rbind(zroundaveragepublic, zroundaverageprivate)

barplot(zrbr,
        names.arg = c("A", "Bk", "Bx", "E", "Gx", "Gh", "Hv", "Hw", "Ht", "M"),
        col = c("lightskyblue2", "mistyrose2"),
        main = "Mean Z Score by Tournament",
        ylab = "Mean Z Score",
        ylim = c(27, 29),
        xpd = FALSE,
        cex.axis = 1,
        cex.main = 1,
        beside = TRUE)
legend("bottomright", legend = c("Public", "Private"), col = c("lightskyblue2", "mistyrose2"), pch = 19, bg = "white")
```



```{r echo=FALSE}
all$zaverage <- ifelse(all$tournament == "Bronx" | all$tournament == "Glenbrooks" | all$tournament == "Harvard", all$Z/7, all$Z/6)

private <- subset(all, School.Type == "Private")
public <- subset(all, School.Type == "Public")
```
Looking at the mean per round, the private scores are now higher. This shows that round size at tournaments acted as confounding variables when determining public vs. private school competitive success rates. Again, we should find the round average for each individual to use in further calculations so we do not weigh certain tournaments as equal to other tournaments even though they had different sample sizes. The public schools' round average is `r mean(public$zaverage)` and the private schools' round average is `r mean(private$zaverage)`.

```{r echo=FALSE}
# X Bar
zxbar_diff <- mean(private$zaverage) - mean(public$zaverage)

# Standard Error
zse_diff <- sqrt((var(private$zaverage)/length(private$zaverage)) + (var(public$zaverage)/length(public$zaverage)))
```

```{r echo=FALSE}
# 95% Confidence Interval
z95 <- c(zxbar_diff - qnorm(0.975) * zse_diff, zxbar_diff + qnorm(0.975) * zse_diff)

# 99% Confidence Interval
z99 <- c(zxbar_diff - qnorm(0.995) * zse_diff, zxbar_diff + qnorm(0.995) * zse_diff)
```
**Confidence Interval Testing:** Once again, we will look for the 95% and the 99% difference in means confidence intervals and test if 0 is a value within the interval.  
95% Confidence Interval: `r z95` and 99% Confidence Interval: `r z99`  
0 is not a value in this confidence interval so we know the difference between public and private school z-score means is significant.

**Conclusion:** When we found the confidence intervals for our data, zero was never a value in the 95% or the 99% difference in means confidence intervals which indicate that there was a significant difference between the two groups. This difference could be attributed to a number of factors but the most likely explanation is that private schools simply have access to more resources than public schools.  
Limitations: A few limitations is that we have a relatively small sample size because we could only use tournament data from 2020-2021. This was still preferable to using data from other years because COVID influenced much of whether tournaments were online or in-person. It was also better than using other tournaments because other non-octa bid tournaments have different levels of competitiveness, different proportions of turnout, and different standards for judging.
Interval Validity: High, the only possible confounder would be participation (ie. whether speaker scores could be higher for public schools if the proportion of students from all public schools in the United States was the same as the proportion of students from all private schools in the United States).
External Validity: Relatively high, in general, observational studies are relatively high. The only possible confounder could be that we chose octa bid tournaments which definitely have higher private school turnout than local tournaments.


**Future Steps:** Looking at Public School Success by State Funding
```{r echo=FALSE, fig5, fig.align='center', fig.height=3.5, fig.width= 5.5}
statefunding <- export(statefunding,
                       "https://teaching-certification.com/teaching/education-spending-by-state/")


statefunding <- statefunding[c('State', 'Spending.per.pupil....s.')]

for(i in 1:nrow(public)){
  for(j in 1:nrow(statefunding)){
    if(public$State[i] == statefunding$State[j]){
      public$funding[i] <- statefunding$Spending.per.pupil....s.[j]
    }
  }
}

public$funding <- sub('.','', public$funding)
public$funding <- sub(',','', public$funding)
public$funding <- as.numeric(public$funding )

plot(public$funding, public$zaverage, xlim = c(7500, 22000), ylim = c(25, 30), xlab = "State Funding per Student", ylab = "Z-Score Round Average")
lm(public$zaverage~public$funding)

```

In order to further analyze the correlation between access to resources and competitive success, we ran a basic linear regression on data with state funding and Z-score round averages in public schools and found that for each $1000 per student that a state puts into education, Z-score round averages increase by 0.008983. Further analysis on the correlation between distribution of funding by county or region and public school competitive success could help identify what variables are contributing to the success of some public schools but not others. 






